% Copyright (c) 2014 Adobe Systems Incorporated. All rights reserved.

% Licensed under the Apache License, Version 2.0 (the "License");
% you may not use this file except in compliance with the License.
% You may obtain a copy of the License at

% http://www.apache.org/licenses/LICENSE-2.0

% Unless required by applicable law or agreed to in writing, software
% distributed under the License is distributed on an "AS IS" BASIS,
% WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
% See the License for the specific language governing permissions and
% limitations under the License.

\input{frontmatter.tex}

\title{White Paper: Introducing the ActionScript\textsuperscript{\textregistered} 4
Language} \author{Bernd Mathiske, Avik Chaudhuri, Krzysztof Palacz, and Basil Hosmer\\
\\
{\it Adobe Systems Incorporated} \\
{\tt AS4@adobe.com}}
%\date{}                                           % Activate to display a given date or no date

\begin{document}

\maketitle

\pagebreak

\copyright{} 2014 Adobe Systems Incorporated. All rights reserved. %[update year of publication]

{\bf Adobe ActionScript 4 Language White Paper Version 1.0}

This document is protected under copyright law, furnished for informational use
only, is subject to change without notice, and should not be construed as a commitment by Adobe Systems Incorporated.
Adobe Systems Incorporated assumes no responsibility or liability for any errors or inaccuracies
that may appear in the informational content contained in this guide.

This guide contains links to third-party websites that are not under the control of Adobe Systems
Incorporated, and Adobe Systems Incorporated is not responsible for the content on any linked site.
If you access a third-party website mentioned in this guide, then you do so at your own risk. Adobe
Systems Incorporated provides these links only as a convenience, and the inclusion of the link does
not imply that Adobe Systems Incorporated endorses or accepts any responsibility for the content on
those third-party sites. No right, license, or interest is granted in any third party technology
referenced in this guide.

This user guide is licensed for use under the terms of the Creative Commons Attribution
Non-Commercial 3.0 License. This License allows users to copy, distribute, and transmit the user
guide for noncommercial purposes only so long as (1) proper attribution to Adobe is given as the
owner of the user guide; and (2) any reuse or distribution of the user guide contains a notice that
use of the user guide is governed by these terms. The best way to provide notice is to include the
following link. To view a copy of this license, visit http://creativecommons.org/licenses/by-nc-sa/3.0/.

Adobe, Adobe AIR, ActionScript, and Flash are either registered
trademarks or trademarks of Adobe Systems Incorporated in the United States and/or other countries. All other trademarks are the
property of their respective owners.  No right or license is granted to any Adobe trademark.
%Updated Information/Additional Third Party Code Information
%available at http://www.adobe.com/go/thirdparty.

Adobe Systems Incorporated, 345 Park Avenue, San Jose, California 95110, USA.

\emph{Notice to U.S. Government End Users:} The Software and Documentation are ``Commercial Items,'' as
that term is defined at 48 C.F.R. \S2.101, consisting of ``Commercial Computer Software'' and
``Commercial Computer Software Documentation,'' as such terms are used in 48 C.F.R. \S12.212
or 48 C.F.R. \S227.7202, as applicable. Consistent with 48 C.F.R. \S12.212 or 48 C.F.R.
\S\S227.7202-1 through 227.7202-4, as applicable, the Commercial Computer Software and
Commercial Computer Software Documentation are being licensed to U.S. Government end users (a) only
as Commercial Items and (b) with only those rights as are granted to all other end users pursuant
to the terms and conditions herein. Unpublished-rights reserved under the copyright laws of the
United States. Adobe agrees to comply with all applicable equal opportunity
laws including, if appropriate, the provisions of Executive Order 11246, as amended, Section 402 of the Vietnam Era
Veterans Readjustment Assistance Act of 1974 (38 USC 4212), and Section 503 of the Rehabilitation
Act of 1973, as amended, and the regulations at 41 CFR Parts 60-1 through 60-60, 60-250, and 60-741.
The affirmative action clause and regulations contained in the preceding
sentence shall be incorporated by reference.

\pagebreak

\tableofcontents

\pagebreak

\input{intro.tex}
\input{dropped.tex}
\input{changes.tex}
\input{new.tex}
\input{outlook.tex}

\appendix
\input{same.tex}
%\input{appendix.tex}

\end{document}
